Enriching Object Detection with 2D-3D Registration and Continuous Viewpoint Estimation

### 2015-04-10 experiments.tex
CVPR cam-ready. reduced the size of images, major updates on the extended cam-ready
cam ready final

\paragraph{Setup.}
We use established benchmark datasets to validate our approach, namely
the 3D Object Classes dataset~\cite{Savarese07}, and
PASCAL3D+~\cite{Xiang14}, a recently proposed extension of Pascal
VOC'12~\cite{pascal12} that provides additional annotations in the
form of aligned 3D CAD models. In both cases, we use the test data
provided by the respective datasets, but train our models entirely
from rendered 3D CAD model images.

@@

\paragraph{Results.}
In Tab.~\ref{tab:who_initializations}, we observe that, on average,
calibration indeed improves performance in terms of AP, sometimes
drastically (e.g., from $54.4$ to $92.8$ for WHO-CG-Z), as observed in
prior work~\cite{Aubry14}. At the same time, calibration is
computationally expensive, resulting in generation time that are two orders of
magnitudes larger than without calibration. Strikingly, our method
NZ-WHO-CG achieves the second best AP of $90.0$ while completing in 79
ms vs. 8.5 s when using calibration.

@@

\begin{table}[!htbp]
  \footnotesize
  \setlength{\tabcolsep}{1pt}
  \centering
  \begin{tabular}{|c|c|r|c|r|}
    \hline
    Methods (AP/MPPE) & before calib.  & time & after calib.~\cite{Aubry14} & time \\
    \hline\hline
    HOG\cite{Dalal05}      & 72.3 / 65.0          &  31ms  & 60.4 / 50.2           & 8.7 sec \\ 
    WHO\cite{Hariharan12}  & 82.1 / \textbf{85.4} &  6162ms& 84.4 / 83.0           & 15.4 sec\\
    WHO-CG                 & 81.7 / 84.9          &  104ms & 83.7 / \textbf{87.3}  & 8.3 sec \\
    WHO-CG-Z               & 54.4 / 65.1          &  103ms & \textbf{92.8} / 86.7  & 8.7 sec \\
    % NZC-WHO-Z    & 89.10 /\textbf{78.64} &    & 91.15 / 74.79                  &     \\
    NZ-WHO-CG {\em (ours)} & \textbf{90.0} / 82.8 &   79ms & 90.3 / 86.8           & 8.5 sec \\
    \hline
  \end{tabular}
  \caption{Average Precision (AP), Mean Precision in Pose Estimation
    (MPPE)~\cite{Lopez-Sastre11}, and generation time for variations of WHO on
    3D Object Classes cars~\cite{Savarese07}. Please see text for
    details.}
  % WHO refers to standard WHO using the
  % method presented in \cite{Hariharan12}, WHO-CG uses iterative
  % Conjugate Gradient method to generate WHO. WHO-CG-Z uses whiten the
  % whole template and zero out textureless region. NZ-WHO-CG is the
  % NZ-WHO which whitens only non-zero cells using iterative Conjugate
  % Gradient method. The time column indicates the time to generate one
  % template. We followed calibration procedure presented in \cite{Aubry14}.}
  \label{tab:who_initializations}
\end{table}

@@

\paragraph{Results.}
Tab.~\ref{tab:3dobject} gives the corresponding results, comparing our
method to two recent state-of-the-art baselines, the aspect layout
model (ALM~\cite{Xiang12}), and the
DPM-VOC+VP~\cite{Pepik12}. Fig.~\ref{fig:3dobject_vis} gives
qualitative results.

@@

To show such ability, we evaluate our method on the PASCAL3D+
dataset~\cite{Xiang14}. This dataset augments PASCAL 2012 images with
high quality viewpoint annotations thus is ideal to measure pose
estimation. The dataset proposes a new metric called Average Viewpoint
Precision (AVP) where it measures the area under viewpoint precision
and detection recall curve. The viewpoint is measured by azimuth
similarity. If the distance between predicted azimuth and ground truth
azimuth is below a certain threshold, the viewpoint is correct.
The baseline methods V-DPM~\cite{Xiang14} and
DPM-VOC+VP~\cite{Pepik12} reported on the PASCAL3D+ dataset are
variants of DPM~\cite{Felzenszwalb10} where each component of DPM
accounts for azimuth. Thus V-DPM and DPM-VOC+VP provide discrete
azimuths only whereas our method provides 3D viewpoint (yaw, pitch,
roll), CAD model instance (model index, rendering, depth) and focal length.
%
\begin{figure}[t]
 \begin{center}
    \setlength\tabcolsep{0pt}
    \begin{tabular}{ccc}
    % \includegraphics[width=0.9\linewidth]{tuning} 
   \includegraphics[width=0.33\linewidth]{tuning/1.png} &
   \includegraphics[width=0.33\linewidth]{tuning/2.png} &
   \includegraphics[width=0.33\linewidth]{tuning/3.png} \\[-5pt]
%   (a) & (b) & (c)\\
   \includegraphics[width=0.33\linewidth]{tuning/4.png} &
   \includegraphics[width=0.33\linewidth]{tuning/5.png} &
   \includegraphics[width=0.33\linewidth]{tuning/6.png} \\[-5pt]
%   (d) & (e) & (f)\\
   \end{tabular}
 \end{center}
 \caption{Effect of fine tuning. (left) original image, (middle) initial detection, (right) continuous fine tuning using Single-Component Metropolis Hastings}
 \label{fig:tuning}
\end{figure}

\paragraph{Setup.}
We use detection bounding boxes provided by both object detectors and
 use our method to perform fine-grained viewpoint estimation. We use
detection bounding boxes from two different methods:
DPM-VOC+VP~\cite{Pepik12} and R-CNN~\cite{Girshick14}, both in their
variants trained on 8 viewpoint categories, since these perform best
in terms of AP. For both cases,
we use the original score for plotting precision-recall curves
(meaning that we can not improve over their AP). We revert back to a default
viewpoint prediction ($0$ angles) in case the confidence of our
method falls below a threshold.

@@

Quantitative results are given in Tab.~\ref{tab:pascal12} in terms of
AP and AVP and in Fig.~\ref{fig:car_cnn_ap} in terms of
precision-recall plots and viewpoint confusion matrices. Example
outputs of our method applied to candidate object detections are given
in Fig.~\ref{fig:pascal12cnn}, and the effect of fine-tuning is
visualized in Fig.~\ref{fig:tuning}. More qualitative results can be found in the
supplemental material.
\paragraph{Results.}
%
In Tab.~\ref{tab:pascal12}, we make two main observations. First, we
see that adding our method to DPM-VOC+VP consistently improves
performance in terms of AVP, for both car and bicycle, and across all
viewpoint bins. This is already the case for our discrete method: the
improvement ranges from $0.3$ for bicycle-8v to $5.8$ for car-4v. 
%
Using the R-CNN as the base detector increases AVP even more, in
particular for the bicycle class: for bicycle-4v, our discrete method
improves over the corresponding DPM-VOC+VP result by $12.6$.

@@

\begin{tabular}{|ccccc|}

@@

\caption{Example enriched bounding boxes. Given R-CNN~\cite{Girshick14}
  detection bounding boxes, our method predicted 2D-3D matching
  reasonably. The first column shows bounding boxes produced by R-CNN detection. Subsequent columns show the output of our method given this bounding box input. Blue boxes are R-CNN output and purple boxes are
  the tightest bounding box enclosing predicted CAD model.}

@@

\begin{figure}[h]
\setlength\tabcolsep{1pt}
\centering
\begin{tabular}{cc}
%  \hline
  \includegraphics[width=0.49\linewidth]{car_cnn4_crop.png} &   
  \includegraphics[width=0.49\linewidth]{car_cnn8_crop.png} \\   
  \includegraphics[width=0.49\linewidth]{car_cnn16_crop.png} &   
  \includegraphics[width=0.49\linewidth]{car_cnn24_crop.png} \\   
%  \hline
\end{tabular}
\caption{Average Precision (AP)(red) and Average Viewpoint Precision (AVP)(green)
  and viewpoint confusion table on PASCAL 12 car validation set using
  R-CNN + Ours (full). From the top left, 4 views, 8 views, 16 views
  and 24 views.}
 % Since our method only augment the object detection,
 %  the AP remains the same for all cases. As we increase the viewpoint
 %  estimation, the AVP decreases. Note that since we estimate null
 %  viewpoint if our method fails to find match, there are more 0
 %  viewpoint predictions (top row of confusion matrices) }
  \label{fig:car_cnn_ap}
\end{figure}

\paragraph{Robustness.}
% Generic object detectors give overlapping detection bounding boxes of the same object which can be handled by non maximal suppression. For these overlapping detections, 
While the R-CNN detector~\cite{Girshick14} is highly robust to
variations in object appearance and even occlusion and truncation,
the resulting bounding box detections vary largely in the object
portions that they contain, which provides a major challenge to any
method that uses these detections as an input for further processing,
such as ours.
%
In order to accommodate this variability, we add a considerable context
region around the proposed bounding boxes before running our method.
We assume that the object can be arbitrarily truncated by the bounding
box and search all plausible scales and translations. This can be
efficiently computed using FFT-based convolution
(Section. \ref{sec:fft}).

>>>>

\vspace{-0.1in}
\paragraph{Setup.}
We use established benchmark datasets to validate our approach, namely
the 3D Object Classes dataset~\cite{Savarese07}, and
PASCAL3D+~\cite{Xiang14}, a recently proposed extension of Pascal
VOC'12~\cite{pascal12} that provides additional annotations in the
form of aligned 3D CAD models. In both cases, we use the test data
provided by the respective datasets, but train our models entirely
from rendered 3D CAD model images.

@@

\vspace{-0.1in}
\paragraph{Results.}
In Tab.~\ref{tab:who_initializations}, we observe that, on average,
calibration indeed improves performance in terms of AP, sometimes
drastically (e.g., from $54.4$ to $92.8$ for WHO-CG-Z), as observed in
prior work~\cite{Aubry14}. At the same time, calibration is
computationally expensive, resulting in generation time that are two orders of
magnitudes larger than without calibration. Strikingly, our method
NZ-WHO-CG achieves the second best AP of $90.0$ while completing in 79
ms vs. 8.5 s when using calibration.

@@

\begin{table}[!htbp]
  \footnotesize
  \setlength{\tabcolsep}{1pt}
  \centering
  \begin{tabular}{|c|c|r|c|r|}
    \hline
    Methods (AP/MPPE) & before calib.  & synth. time & after calib.~\cite{Aubry14} & calib. time \\
    \hline\hline
    HOG\cite{Dalal05}      & 72.3 / 65.0          &  31ms  & 60.4 / 50.2           & 8.7 sec \\ 
    WHO\cite{Hariharan12}  & 82.1 / \textbf{85.4} &  6162ms& 84.4 / 83.0           & 15.4 sec\\
    WHO-CG                 & 81.7 / 84.9          &  104ms & 83.7 / \textbf{87.3}  & 8.3 sec \\
    WHO-CG-Z               & 54.4 / 65.1          &  103ms & \textbf{92.8} / 86.7  & 8.7 sec \\
    % NZC-WHO-Z    & 89.10 /\textbf{78.64} &    & 91.15 / 74.79                  &     \\
    NZ-WHO-CG {\em (ours)} & \textbf{90.0} / 82.8 &   79ms & 90.3 / 86.8           & 8.5 sec \\
    \hline
  \end{tabular}
  \caption{Average Precision (AP), Mean Precision in Pose Estimation
      (MPPE)~\cite{Lopez-Sastre11} of variants of WHO on 3D Object Classes
      cars~\cite{Savarese07}, and their corresponding synthesis and calibration
      time per template. Please see text for details.}
  % WHO refers to standard WHO using the
  % method presented in \cite{Hariharan12}, WHO-CG uses iterative
  % Conjugate Gradient method to generate WHO. WHO-CG-Z uses whiten the
  % whole template and zero out textureless region. NZ-WHO-CG is the
  % NZ-WHO which whitens only non-zero cells using iterative Conjugate
  % Gradient method. The time column indicates the time to generate one
  % template. We followed calibration procedure presented in \cite{Aubry14}.}
  \label{tab:who_initializations}
\end{table}

@@

\vspace{-0.1in}
\paragraph{Results.}
Tab.~\ref{tab:3dobject} gives the corresponding results, comparing our
method to two recent state-of-the-art baselines, the aspect layout
model (ALM~\cite{Xiang12}), and the
DPM-VOC+VP~\cite{Pepik12}. Fig.~\ref{fig:3dobject_vis} gives
qualitative results.

@@

To show such ability, we evaluate our method on the PASCAL3D+
dataset~\cite{Xiang14}. This dataset augments PASCAL 2012 images with
high quality viewpoint annotations thus is ideal to measure pose
estimation. The dataset proposes a new metric called Average Viewpoint
Precision (AVP) where it measures the area under viewpoint precision
and detection recall curve. The viewpoint is measured by azimuth
similarity. If the distance between predicted azimuth and ground truth
azimuth is below a certain threshold, the viewpoint is correct.
The baseline methods V-DPM~\cite{Xiang14} and
DPM-VOC+VP~\cite{Pepik12} reported on the PASCAL3D+ dataset are
variants of DPM~\cite{Felzenszwalb10} where each component of DPM
accounts for azimuth. Thus V-DPM and DPM-VOC+VP provide discrete
azimuths only whereas our method provides 3D viewpoint (yaw, pitch,
roll), CAD model instance (model index, rendering, depth) and focal length.
%
\begin{figure}[t]
 \begin{center}
    \setlength\tabcolsep{0pt}
    \begin{tabular}{ccc}
    % \includegraphics[width=0.9\linewidth]{tuning} 
   \includegraphics[width=0.33\linewidth]{tuning/1_sm.png} &
   \includegraphics[width=0.33\linewidth]{tuning/2_sm.png} &
   \includegraphics[width=0.33\linewidth]{tuning/3_sm.png} \\[-5pt]
%   (a) & (b) & (c)\\
   \includegraphics[width=0.33\linewidth]{tuning/4_sm.png} &
   \includegraphics[width=0.33\linewidth]{tuning/5_sm.png} &
   \includegraphics[width=0.33\linewidth]{tuning/6_sm.png} \\[-5pt]
%   (d) & (e) & (f)\\
   \end{tabular}
 \end{center}
 \caption{Effect of fine tuning. (left) original image, (middle) initial detection, (right) continuous fine tuning using Single-Component Metropolis Hastings}
 \label{fig:tuning}
\end{figure}

\vspace{-0.1in}
\paragraph{Setup.}
We use detection bounding boxes provided by both object detectors and
 use our method to perform fine-grained viewpoint estimation. We use
detection bounding boxes from two different methods:
DPM-VOC+VP~\cite{Pepik12} and R-CNN~\cite{Girshick14}, both in their
variants trained on 8 viewpoint categories, since these perform best
in terms of AP. For both cases,
we use the original score for plotting precision-recall curves
(meaning that we can not improve over their AP). We revert back to a default
viewpoint prediction ($0^\circ$ azimuth) in case the confidence of our
method falls below a threshold.

@@

Quantitative results are given in Tab.~\ref{tab:pascal12} in terms of
AP and AVP and in Fig.~\ref{fig:car_cnn_ap} in terms of
precision-recall plots and viewpoint confusion matrices. Example
outputs of our method applied to candidate object detections are given
in Fig.~\ref{fig:pascal12cnn}, and the effect of fine-tuning is
visualized in Fig.~\ref{fig:tuning}. More qualitative results can be found in the
supplemental material.

\vspace{-0.1in}
\paragraph{Results.}
%
In Tab.~\ref{tab:pascal12}, we make two main observations. First, we
see that adding our method to DPM-VOC+VP consistently improves
performance in terms of AVP, for both car and bicycle, and across all
viewpoint bins. This is already the case for our discrete method: the
improvement ranges from $0.3$ for bicycle-8v to $5.8$ for car-4v. 
%
Using the R-CNN as the base detector increases AVP even more, in
particular for the bicycle class: for bicycle-4v, our discrete method
improves over the corresponding DPM-VOC+VP result by $12.6$.

@@

\begin{tabular}{|c|c|c|c|c|}

@@

\caption{Examples of enriched bounding boxes. Given R-CNN~\cite{Girshick14}
    detection bounding boxes, our method predicted 2D-3D matching reasonably.
    The first column shows bounding box candidates produced by R-CNN detection.
    Subsequent columns show the output of our method given the bounding box
    from R-CNN detection candidates. Blue boxes are R-CNN output and purple
    boxes are the tightest bounding box enclosing predicted CAD model.}

@@

\begin{figure}[h]
\setlength\tabcolsep{1pt}
\centering
\begin{tabular}{c}
%  \hline
%  \includegraphics[width=0.49\linewidth]{car_cnn4_crop.png}    
%  \includegraphics[width=0.49\linewidth]{car_cnn8_crop.png} \\   
%  \includegraphics[width=0.49\linewidth]{car_cnn16_crop.png} &   
  \includegraphics[width=0.9\linewidth]{car_cnn24_crop.png} \\   
%  \hline
\end{tabular}
\caption{Average Precision (AP)(red) and Average Viewpoint Precision (AVP)(green)
  and viewpoint confusion table on PASCAL 12 car validation set using
  R-CNN + Ours (full) for 24 views. All four viewpoint discretizations are available on the supplementary paper.}
 % Since our method only augment the object detection,
 %  the AP remains the same for all cases. As we increase the viewpoint
 %  estimation, the AVP decreases. Note that since we estimate null
 %  viewpoint if our method fails to find match, there are more 0
 %  viewpoint predictions (top row of confusion matrices) }
  \label{fig:car_cnn_ap}
\end{figure}

\vspace{-0.1in}
\paragraph{Robustness.}
% Generic object detectors give overlapping detection bounding boxes of the same object which can be handled by non maximal suppression. For these overlapping detections, 
While the R-CNN detector~\cite{Girshick14} is highly robust to
variations in object appearance and even occlusion and truncation,
the resulting bounding box detections vary largely in the object
portions that they contain, which provides a major challenge to any
method that uses these detections as an input for further processing,
such as ours.
%
In order to accommodate this variability, we add a considerable context
region around the proposed bounding boxes before running our method.
We assume that the object can be arbitrarily truncated by the bounding
box and search all plausible scales and translations. This can be
efficiently computed using FFT-based convolution
(Section. \ref{sec:fft}).

### 2015-04-10 mcmc.tex
CVPR cam-ready. reduced the size of images, major updates on the extended cam-ready
cam ready final

\paragraph{Inference.}
We approximate the MAP solution for $\theta$ by drawing samples from the
distribution $P(\theta | \mathcal{I})$, using the Metropolis-Hastings
algorithm. Specifically, we use a variant that changes only a single
component of the parameter vector $p$ at a time, termed Single
Component Metropolis-Hastings \cite{mcmc}.

This algorithm changes the current state $p$ to a new state $q$ based
on the acceptance probability
\begin{align}
    A(\theta \rightarrow \theta^+) & =  min\left( 1, \frac{P(\theta^+ |
            \mathcal{I}) g(\theta^+ \rightarrow \theta)}{P(\theta | \mathcal{I})
            g(\theta \rightarrow \theta^+)}\right)
\end{align}
We define $3$ different types of moves that can alter the state
(Fig.~\ref{fig:moves}), {\em (i)} changing the focal length $f$, {\em
(ii)} one of the rotational pose parameters $v_i$, and {\em (iii)} CAD model
index $m$. For (i) and (ii), we use Gaussian proposal distributions, and a
uniform distribution over model changes for (iii). We implicitly compute all
possible translation and scale by convolving $w(\theta^+)$ with a HOG pyramid.
\begin{align}
    g(\theta \rightarrow \theta(v_i^+)) & \sim
    \mathcal{N}(\theta_{v_i},\sigma_v) \quad \mbox{ for }\; i \in \{1,2,3\}\\
    g(\theta \rightarrow \theta(f^+)) & \sim \mathcal{N}(\theta_{f}, \sigma_f)\\
    g(\theta \rightarrow \theta(m^+)) & \sim (1-c) \delta(\theta_m) +
    c\,\textrm{Unif}(1,M)
\end{align}
where $g$ is the proposal distribution, and $m \in \{1,\cdots,M\}$. 
% and $P$ is the distribution of which we
%defined using Gaussian distribution around previous state. $A$ is the
%acceptance probability where we accept new proposal with $A(x \rightarrow x')$.
%
% One can also think of the image as an image cropped by detection
% bounding box. We synthesize covariance matrix and create NZ-WHO
% template $w(x)$ for particular viewpoint, focal length and CAD model.
% In sum, for new proposal $x'$, we make a NZ-WHO template on the fly 
% We found out that it requires good initialization to converge to
% local minima.

@@

% there are several advantages that was not possible such as
% \textbf{on-the-fly} template generation and making large number of
% templates
\begin{figure}[t]
\centering
    \includegraphics[width=0.7\linewidth]{tuning2} \\ [-5pt]
    \caption{Search space that our fine-tuning stage handles (Sect.~\ref{sec:fine}).}
 \label{fig:moves}
\end{figure}

>>>>

\paragraph{Inference.}
We approximate the MAP solution for $\theta$ by drawing samples from the
distribution $P(\theta | \mathcal{I})$, using the Metropolis-Hastings
algorithm. Specifically, we use a variant that changes only a single
component of the parameter vector $\theta$ at a time, termed Single
Component Metropolis-Hastings \cite{mcmc}.

This algorithm changes the current state $\theta$ to a new state $\theta^+$ based
on the acceptance probability
\begin{align}
    A(\theta \rightarrow \theta^+) & =  min\left( 1, \frac{P(\theta^+ |
            \mathcal{I}) g(\theta^+ \rightarrow \theta)}{P(\theta | \mathcal{I})
            g(\theta \rightarrow \theta^+)}\right)
\end{align}
We define $3$ different types of moves that can alter the state
(Fig.~\ref{fig:moves}), {\em (i)} changing the focal length $f$, {\em
(ii)} one of the rotational pose parameters $v_i$, and {\em (iii)} CAD model
index $m$. For (i) and (ii), we use Gaussian proposal distributions, and a
uniform distribution over model changes for (iii). We implicitly compute all
possible translation and scale by convolving $w(\theta^+)$ with a HOG pyramid.
\begin{align}
    g(\theta \rightarrow \theta(v_i^+)) & \sim
    \mathcal{N}(\theta_{v_i},\sigma_v) \quad \mbox{ for }\; i \in \{1,2,3\}\\
    g(\theta \rightarrow \theta(f^+)) & \sim \mathcal{N}(\theta_{f}, \sigma_f)\\
    g(\theta \rightarrow \theta(m^+)) & \sim (1-c) \delta(\theta_m) +
    c\,\textrm{Unif}(1,M)
\end{align}
where $g$ is the proposal distribution, and $m \in \{1,\cdots,M\}$. 
% and $P$ is the distribution of which we
%defined using Gaussian distribution around previous state. $A$ is the
%acceptance probability where we accept new proposal with $A(x \rightarrow x')$.
%
% One can also think of the image as an image cropped by detection
% bounding box. We synthesize covariance matrix and create NZ-WHO
% template $w(x)$ for particular viewpoint, focal length and CAD model.
% In sum, for new proposal $x'$, we make a NZ-WHO template on the fly 
% We found out that it requires good initialization to converge to
% local minima.

@@

% there are several advantages that was not possible such as
% \textbf{on-the-fly} template generation and making large number of
% templates
\begin{figure}[t]
\centering
    \includegraphics[width=0.7\linewidth]{tuning2_small} \\ [-5pt]
    \caption{Search space that our fine-tuning stage handles (Sect.~\ref{sec:fine}).}
 \label{fig:moves}
\end{figure}
